---
title: "2016 beeswarm plots"
author: "Kim Cressman"
date: "`r Sys.Date()`"
output: 
    html_document:
        keep_md: true
        code_folding: 'hide'
        toc: yes
        toc_float: true
        toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```


Open up some libraries:

```{r}
library(SWMPr)  # for easy import and QC of SWMP data
library(tidyverse)  # for lots of data manipulation functions
library(ggbeeswarm)  # for cool plotting
library(knitr)  # for pretty tables
```


# Bring in the data  


```{r}
# look for data here
datapath <- "C:/Users/kimberly.cressman/Desktop/Main Docs/Data-latest"

# import data
dat_bl <- import_local(datapath, "gndblwq2016") %>%
    qaqc(qaqc_keep = c(0, 1, 5)) %>%
    select(-level, -clevel, -chlfluor)

dat_bh <- import_local(datapath, "gndbhwq2016") %>%
    qaqc(qaqc_keep = c(0, 1, 5)) %>%
    select(-level, -clevel, -chlfluor)

dat_bc <- import_local(datapath, "gndbcwq2016") %>%
    qaqc(qaqc_keep = c(0, 1, 5)) %>%
    select(-level, -clevel, -chlfluor)

dat_pc <- import_local(datapath, "gndpcwq2016") %>%
    qaqc(qaqc_keep = c(0, 1, 5)) %>%
    select(-level, -clevel, -chlfluor)
```

<br>

Next, I want to get all of the salinities into a single data frame, with the site name as the column name, so I can easily put them all on a plot and color by site. Surely there's a better way, but here's what I've gotten to work.



# Functions to extract a single parameter

## Wide Format: param_extract_wide

```{r}
#' Wide Parameter Extraction from multiple data frames
#' 
#' This function pulls one specified parameter out of all the water quality data frames, and glues them together into one wide data frame.
#'
#' @param BC data frame with Bayou Cumbest data. if unspecified, defaults to dat_bc.
#' @param BH data frame with Bayou Heron data
#' @param BL data frame with Bangs Lake data
#' @param PC data frame with Point aux Chenes data
#' @param param put in quotes - param you want to extract. examples: "sal" / "do_mgl"
#'
#' @return returns one wide data frame, with column names being site abbreviations (BC, BH, BL, PC) and values being the parameter of interest
#' @export
#'

param_extract_wide <- function(param, BC = dat_bc, BH = dat_bh, BL = dat_bl, PC = dat_pc) {
    bc <- BC %>%
        select(datetimestamp, !!param) %>%
        rename(BC = !!param)
    bh <- BH %>%
        select(datetimestamp, !!param) %>%
        rename(BH = !!param)
    bl <- BL %>%
        select(datetimestamp, !!param) %>%
        rename(BL = !!param)
    pc <- PC %>%
        select(datetimestamp, !!param) %>%
        rename(PC = !!param)
   
    # glue them all together 
    all <- bc %>%
        left_join(bh) %>%
        left_join(bl) %>%
        left_join(pc)
    
    return(all)
        
}
```


## Long format: param_extract_long

```{r}
#' Long Parameter Extraction from multiple data frames
#' 
#' This function pulls one specified parameter out of all the water quality data frames, and glues them together into one long data frame.
#'
#' @param BC data frame with Bayou Cumbest data. if unspecified, defaults to dat_bc.
#' @param BH data frame with Bayou Heron data
#' @param BL data frame with Bangs Lake data
#' @param PC data frame with Point aux Chenes data
#' @param param put in quotes - param you want to extract. examples: "sal" / "do_mgl"
#'
#' @return returns one long data frame, with a column specifying site and a column specifying the value at that site
#' @export
#'

param_extract_long <- function(param, BC = dat_bc, BH = dat_bh, BL = dat_bl, PC = dat_pc) {
    bc <- BC %>%
        select(datetimestamp, !!param) %>%
        rename(BC = !!param)
    bh <- BH %>%
        select(datetimestamp, !!param) %>%
        rename(BH = !!param)
    bl <- BL %>%
        select(datetimestamp, !!param) %>%
        rename(BL = !!param)
    pc <- PC %>%
        select(datetimestamp, !!param) %>%
        rename(PC = !!param)
   
    # glue them all together and gather into a long format 
    all <- bc %>%
        left_join(bh) %>%
        left_join(bl) %>%
        left_join(pc) %>%
        gather(key = "site", value = "value", -datetimestamp) %>%
        mutate(site = factor(site),
               site = fct_relevel(site, "BH", "BC", "BL", "PC"))
    
    return(all)
        
}
```


# Extract parameters

<br>

__Temperature__


```{r}
temp_long <- param_extract_long("temp")
head(temp_long)
tail(temp_long)
```

<br>

__Salinity__

```{r}
sal_long <- param_extract_long("sal")
head(sal_long)
tail(sal_long)
```

<br>

__Dissolved Oxygen mg/L__

```{r}
domgl_long <- param_extract_long("do_mgl")
head(domgl_long)
tail(domgl_long)
```

<br>

__Dissolved Oxygen % saturation__

```{r}
dopct_long <- param_extract_long("do_pct")
head(dopct_long)
tail(dopct_long)
```

<br>

__pH__

```{r}
ph_long <- param_extract_long("ph")
head(ph_long)
tail(ph_long)
```

<br>

__Depth__  

(not corrected for barometric pressure)

```{r}
depth_long <- param_extract_long("depth")
head(depth_long)
tail(depth_long)
```



<br>

# Make plots

```{r}
beeswarm_wq <- function(dat, title = "", xlab = "Site", ylab = "value") {
    p <- ggplot(dat) +
        geom_quasirandom(aes(x = site, y = value, color = site), 
                         na.rm = TRUE, # ignore missing data
                         show.legend = FALSE) +  # don't need a legend for colors by site
        labs(title = title, 
             x = xlab, 
             y = ylab) +
        theme_bw()
    
    print(p)
}
```

## Temperature

```{r}
beeswarm_wq(temp_long, title = "2016 Water Temperature")
```


## Salinity

```{r}
beeswarm_wq(sal_long, title = "2016 Salinity")
```


## DO mg/L

```{r}
beeswarm_wq(domgl_long, title = "2016 DO mg/L")
```

## DO % saturation

```{r}
beeswarm_wq(dopct_long, title = "2016 DO % saturation")
```

## pH

```{r}
beeswarm_wq(ph_long, title = "2016 pH")
```


## depth

```{r}
beeswarm_wq(depth_long, title = "2016 depth")
```
